The transiting dust clumps in the evolved disk of the Sun-like UXor RZ~Psc

### 2016-11-30 ms.tex
revised version sent to co-authors

\title[The transiting dust clumps of RZ~Psc]{The transiting dust clumps in the evolved
  disk of the Sun-like UXOr RZ~Psc}

\author[Grant M. Kennedy et al]{Grant M. Kennedy$^1$,
  Matthew A. Kenworthy$^2$,
  Joshua Pepper$^3$,
  Joseph E. Rodriguez$^{4,5}$, 
  Robert J. Siverd$^6$, \&
  Keivan G. Stassun$^{5,7}$}

@@

\begin{abstract}
  RZ~Psc is a young Sun-like star, long associated with the UXOr class of variable stars,
  that is occulted by optically thick dust clumps several times each year. The system has
  an infrared excess, interpreted as evidence that the dimming events are the passage of
  asteroidal fragments in front of the host star. Here, we present a decade of optical
  photometry of RZ~Psc and take a critical look at the asteroid belt interpretation. We
  find weak evidence for periodicity near 65 days (0.3au), and show that the distribution
  of light curve gradients is non-uniform for deep events, which we interpret as possible
  evidence for an asteroidal fragment-like clump structure. However, the clumps are very
  likely seen above a high optical depth mid-plane, so the disk's bulk clumpiness is not
  revealed. While circumstantial evidence suggests an asteroid belt is more plausible
  than a gas-rich transition disk with an inner hole, the evolutionary status remains
  uncertain. We suggest that the rarity of Sun-like stars showing disk-related
  variability may arise because i) any accretion streams are transparent, and/or ii)
  turbulence above the inner rim is normally shadowed by a flared outer disk.
\end{abstract}

@@

Disks of gas and dust surround essentially all young analogues of our Sun
(e.g. \cite{2001ApJ...553L.153H}). The lifetime of the gas in these disks is very short
compared to the stellar lifetime, and within a few million years most has accreted onto
the star and the remainder been lost to space in photoevaporative flows. The evolution of
the dust during this phase is uncertain, but the existence of gas giant planets makes it
clear that planetary building blocks, and of course some planets, form on a similar or
shorter timescale.

@@

A promising way to probe these inner regions is by observing temporal variability
(e.g. \cite{2014Sci...345.1032M}). Optical and IR stellar variation has been studied
for decades (e.g. \cite{1945ApJ...102..168J,1994AJ....108.1906H}), and has recently
been reinvigorated by large scale efforts
(e.g. \cite{2011ApJ...733...50M,2014AJ....147...82C}) and as a side-effect of
large-scale surveys for transiting planets
\cite{2012AJ....143...72M,2013AJ....146..112R,2016MNRAS.457.3988B}. Of many different
classes of variables, the ones of most interest and relevance here are the
``UXOrs''. These are usually Herbig Ae/Be stars, and typically show several magnitudes of
extinction that is generally attributed to variable obscuration by circumstellar dust
\cite{1994AJ....108.1906H,1999AJ....118.1043H,2000A&A...364..633N}.\footnote{Not all
  stars occulted by dust are UXOrs. Two main classes are: those occulted by circumstellar
  material beyond $\gtrsim$10au, such as AA~Tau and V409~Tau which show
  $\gtrsim$year-long dimming events \cite{2013A&A...557A..77B,2015AJ....150...32R}, and
  systems such as $\epsilon$~Aur, J1407, EE~Cep, and OGLE-LMC-ECL-11893 where the
  occultations are attributed to circumsecondary disks
  (e.g. \cite{1999MNRAS.303..521M,2012AJ....143...72M,2014ApJ...788...41D,2015ApJS..220...14K}).}
Two key arguments that favour circumstellar dust as the cause are i) ``blueing'', where
the star is reddened when moderately dimmed but returns to the stellar colour for very
deep ($\gtrsim$1mag) events -- the remaining flux is interpreted as light scattered off
circumstellar dust \cite{1988SvAL...14...27G}, and ii) increased polarisation fraction
during dimming events, caused by a greater fraction of the flux being contributed by
dust-scattered light (e.g. \cite{2001ARep...45...51R}). UXOrs therefore reveal
information on the degree of non-axisymmetry, the ``clumpiness'', of dust orbiting a star
on spatial scales similar to the star itself. The observations can span multiple orbits
to test for repeated dimming events (e.g. \cite{1999A&A...349..619B}), and by using
different bandpasses and polarisation can estimate dust grain sizes
(e.g. \cite{1988SvAL...14...27G,1994ASPC...62...63G}).

In the majority of UXOr-like cases (i.e. those related to obscuration by dust), including
other classes such as ``dippers''
(e.g. \cite{2014AJ....147...82C,2016ApJ...816...69A}), the processes causing young
stars to vary are attributed to gas-rich protoplanetary disks. For Herbig Ae/Be stars the
obscuration is thought to be caused by hydrodynamic turbulence that lifts dust above the
puffed up inner rim of a self-shadowed disk \cite{2003ApJ...594L..47D}. For the dippers,
which are observed around low-mass stars, the obscuration is attributed to dust in
accretion streams that link the inner disk and the stellar surface, and/or to variations
in the height of the inner disk edge
\cite{1999A&A...349..619B,2016ApJ...816...69A,2016arXiv160503985B}. The common theme is
therefore that the location of the occulting dust is as close to the star as physically
possible, being set by sublimation. These systems tell us about the nature of turbulence
and accretion in gas-dominated disks, but so far reveal little about how these disks
transition to the debris phase and the subsequent evolution.

Here we focus on RZ~Psc, a star that shows UXOr-like variability
(e.g. \cite{1985PZ.....22..181Z,1999A&AS..140..293G,2003ARep...47..580S}). As a young
K0V type star with no evidence for gas accretion, but a strong infrared (IR) excess
suggestive of an Asteroid belt analogue, this system appears unique among UXOrs and may
provide new information on the structure of inner planetary systems following dispersal
of the gas disk \cite{2010A&A...524A...8G,2013A&A...553L...1D}. Specifically, we use a
decade of ground-based optical photometry of RZ~Psc (section \ref{s:data}) to draw
conclusions on dust location (section \ref{s:where}), and discuss the possible disk
structure and evolutionary state in section \ref{s:disk}. We conclude in section
\ref{s:conc}.

@@

There is significant evidence that the optical variations seen towards RZ~Psc are caused
by circumstellar dust: i) during dimming events the colour becomes redder
\cite{1985PZ.....22..181Z,1980PZ.....21..310K} in a way consistent with that expected for
dust \cite{1981Afz....17...87P,2004ARep...48..470P}, ii) the maximum depth is about
2.5mag and during these events the colour returns to near stellar values, suggesting that
the remaining emission is from light scattered off the circumstellar dust (i.e. the star
is fully occulted \cite{1981Afz....17...87P,1988SvAL...14...27G}, and iii) the
polarisation fraction increases during the transits, as expected if an increasing
fraction of the light is scattered off circumstellar dust
\cite{1988SvAL...14...27G,1991Afz....34..333K,2003ARep...47..580S}.

What separates RZ~Psc from other UXOrs (and dippers) is i) the spectral type is K0V
rather than Herbig Ae/Be for UXOrs and late K to M-type for dippers, ii) the occulting
dust lies well beyond the sublimation radius, and iii) the star is not associated with a
star-forming region so is inferred to be a few tens of millions of years old. The dust
distance has been inferred from the speed of ingress of dimming events, which was
previously estimated as about 0.6au (for circular orbits
\cite{2013A&A...553L...1D}). Corroborating evidence comes from the $\sim$500K temperature
of the dust seen in the mid-IR, which places it near 0.4-0.7au (depending on optical
depth) and therefore at a location consistent with the occulting dust
\cite{2013A&A...553L...1D}. The distance to RZ~Psc is unknown, but as an apparently
isolated star that shows Li absorption the age has been estimated as a few tens of Myr,
and therefore beyond the age at which a gas-rich disk should exist
\cite{2010A&A...524A...8G,2014A&A...563A.139P}. Further distinguishing features are that
the duration of the dimming events is relatively short compared to other UXOrs, a few
days rather than a few weeks, and that no near-IR excess or accretion signatures are seen
\cite{2014A&A...563A.139P}, so interpretations related to accretion of disk material onto
the star (e.g. \cite{1999AJ....118.1043H,1999A&A...349..619B,2016arXiv160503985B}) are
unlikely.

@@

@@

% yearly-lc.pro
\begin{figure*}
  \begin{center}
    \hspace{-0.5cm} \includegraphics[width=\textwidth]{figs/yearly-2004.eps}
    \caption{WASP and KELT-North data.}\label{fig:waspkelt}
  \end{center}
\end{figure*}

To study the temporal variability of RZ~Psc we use two seasons of public data from the
Wide-Angle Search for Planets (WASP \cite{2006PASP..118.1407P}), and nine seasons of data
from the Kilodegree Extremely Little Telescope North (KELT-North
\cite{2007PASP..119..923P}). We also collected, but ultimately did not use, photometric
observations of RZ~Psc from a wide variety of other sources
(\cite{1994AJ....108.1906H,1973IBVS..783....1K,1980PZ.....21..310K,1985PZ.....22..181Z,1991Afz....34..333K,1997AcA....47..467P,2014Ap.....57..491P},
the Catalina Sky Survey, the American Association of Variable Star Observers, the All-Sky
Automated Survey). Most of these other data have been previously published and their
collation here is largely for the benefit of others. This compilation contains nearly all
data for which tables have been published, but aside from the Harvard plate photometry
published by G\:urtler et al. \cite{1999A&AS..140..293G} we have not sought other
photometry so the light curve remains incomplete.\footnote{Normalised photometry is
  available at \href{https://github.com/drgmk/rzpsc}{https://github.com/drgmk/rzpsc}}

Here we focus on the WASP and KELT-North data, as it has not been previously analysed and
has considerably higher cadence (many measurements per night) and temporal coverage
(nightly, weather permitting) than other data sets. The WASP data from 2004 and 2006 are
public and were obtained from an online
archive\footnote{\href{http://wasp.cerit-sc.cz}{http://wasp.cerit-sc.cz}}. These data
were processed in a manner similar to that described by \cite{2014MNRAS.441.2845V},
where common-mode variations were removed using 50 quiet nearby stars. The KELT-North
data, 2006-2014, were used in raw form, the only specific treatment being a 4\% relative
correction being made to ensure observations taken in the ``east'' and ``west'' telescope
orientations have the same calibration. For a full description of the KELT-North data
reduction, see \cite{2012ApJ...761..123S}.

We normalised each year's data from each instrument separately by converting magnitudes
to flux density and dividing out the sigma-clipped median. In doing so we are assuming
that variations due to the slightly different filter bandpasses are unimportant. We
normalise the light curve to have an out-of-occultation baseline of 1 (Figure
\ref{fig:waspkelt}). Each row shows a season's data, starting on May 1 each year (JD also
indicated). Most year's data therefore extend into the next year, so the ``2006 data''
refers to data from the 2006/2007 observing season.

\subsection{Qualitative light curve overview}\label{ss:quallc}

@@

It is clear from Figure \ref{fig:waspkelt} that RZ~Psc undergoes the very deep dimming
events that are typical of UXOrs. These are seen a few times each observing season and
vary in complexity, with a few extended events (e.g. 2006) and a greater number of
``neater'' single events (e.g. 2010). In some years there is also significant variability
at shallower depths. Of particular note is the pair of deep events in 2006; these appear
to be about 70 days apart, and given the suggestion that the asteroid belt analogue
resides near 0.5au a natural inference is that these two events are related. If true,
this repetition corresponds to a semi-major axis of about 0.3au, which given
uncertainties in the true disk spectrum could be consistent with the location of the
asteroid belt. In 2004 there are about 100 days of near-consecutive nights of data and no
deep events, so either the true period is longer than 100 days, or dust clumps can be
created (and perhaps destroyed) on timescales of a year or so.

@@

@@

The main part of our analysis concerns attempts to extract information from the light
curves, taking advantage of the great number of dimming events seen over ten seasons. In
this section we focus on the radial location of the bodies (``clumps'') that pass in
front of RZ~Psc, first searching for periodicity associated with repeat events, and then
using the light curve gradients to constrain the projected velocities. This analysis
primarily focusses on what can be gleaned from the light curves, and the implications of
these results for different clump origins are then explored in section \ref{s:disk}.

@@

We begin by estimating the lifetime of an occulting clump as a check on the plausibility
that dimming events should repeat. The rate at which clumps are sheared out is
$dv/dr = - \Omega/2$, which accounting for shear in both forward (interior) and backward
(exterior) directions is to first order ($R_{\rm cl} \ll a$)
\begin{equation}\label{eq:shear}
  v_{\rm sh} = R_{\rm cl} \Omega \, ,
\end{equation}
so the clump expansion rate due to shear in units of clump radii is the same as the
orbital frequency. That is, for an orbital period of a year, after one orbit a clump will
be stretched by a factor of $2\pi$ around the orbit, and the optical depth will be $2\pi$
lower (though it may still be optically thick). Thus, clumps that are not gravitationally
bound are expected to have a short lifetime at optical depths that are large enough to
cause detectable dimming events, but could cause repeated dimming events if they are
initially optically thick.

@@

% period-search.pro
\begin{figure}
\begin{center}
\hspace{-0.5cm} \includegraphics[width=0.7\textwidth]{figs/auto.eps}
\caption{Discrete autocorrelation function for yearly WASP and KELT-North data, computed for
  lags between 10-155 days). The second and third lines from the top show all data, and
  2006-excluded data. Squares show points more than 3$\sigma$ from the sigma-clipped
  mean. The topmost line shows the number of years that show a peak (square) within each
  5 day bin. The peak at 70 days is 7 years and the dashed line is zero.}\label{fig:auto}
\end{center}
\end{figure}

>>>>

\title[The transiting dust clumps of RZ~Psc]{The transiting dust clumps in the evolved
  disk of the Sun-like UXor RZ~Psc}

\author[Grant M. Kennedy et al]{Grant M. Kennedy$^1$,
  Matthew A. Kenworthy$^2$,
  Joshua Pepper$^3$,
  Joseph E. Rodriguez$^{4,5}$, 
  Robert J. Siverd$^6$,
  Keivan G. Stassun$^{5,7}$, \&
  Mark C. Wyatt$^1$}

@@

\begin{abstract}
  RZ~Psc is a young Sun-like star, long associated with the UXor class of variable stars,
  which is partially or wholly dimmed by dust clumps several times each year. The system
  has a bright and variable infrared excess, which has been interpreted as evidence that
  the dimming events are the passage of asteroidal fragments in front of the host
  star. Here, we present a decade of optical photometry of RZ~Psc and take a critical
  look at the asteroid belt interpretation. We show that the distribution of light curve
  gradients is non-uniform for deep events, which we interpret as possible evidence for
  an asteroidal fragment-like clump structure. However, the clumps are very likely seen
  above a high optical depth mid-plane, so the disk's bulk clumpiness is not
  revealed. While circumstantial evidence suggests an asteroid belt is more plausible
  than a gas-rich transition disk, the evolutionary status remains uncertain. We suggest
  that the rarity of Sun-like stars showing disk-related variability may arise because i)
  any accretion streams are transparent, and/or ii) turbulence above the inner rim is
  normally shadowed by a flared outer disk.
\end{abstract}

@@

Disks of gas and dust surround essentially all young analogues of our Sun
(e.g. \cite{2001ApJ...553L.153H}). The lifetime of the gas in these disks is very short
compared to the stellar lifetime, and within a few million years has accreted onto the
star, been lost to space in photoevaporative flows, and contributed to building
planets. The evolution of the dust during this phase is uncertain, but the existence of
gas giant planets makes it clear that planetary building blocks, and of course some
planets, form on a similar or shorter timescale.

@@

A promising way to probe these inner regions is by observing temporal variability
(e.g. \cite{2014Sci...345.1032M}). Optical and IR stellar variation has been studied for
decades (e.g. \cite{1945ApJ...102..168J,1994AJ....108.1906H}), and has recently been
reinvigorated by large scale efforts
(e.g. \cite{2011ApJ...733...50M,2014AJ....147...82C}) and as a side-effect of large-scale
surveys for transiting planets \cite{2012AJ....143...72M,2013AJ....146..112R}. Of many
different classes of variables, the ones of most interest and relevance here are the
``UXors'', named for the prototypical system UX~Orionis \cite{1994AJ....108.1906H}. These
are usually Herbig Ae and late-type Herbig Be stars \cite{1984A&AS...55..109F}, and
typically show several magnitudes of extinction that is generally attributed to variable
obscuration by circumstellar dust
\cite{1994AJ....108.1906H,1999AJ....118.1043H,2000A&A...364..633N}.\footnote{Not all
  stars occulted by dust are UXors. Two main classes are: those occulted by circumstellar
  material beyond $\gtrsim$10au, such as AA~Tau and V409~Tau which show
  $\gtrsim$year-long dimming events \cite{2013A&A...557A..77B,2015AJ....150...32R}, and
  systems such as $\epsilon$~Aur, J1407, EE~Cep, and OGLE-LMC-ECL-11893 where the
  occultations are attributed to circumsecondary disks
  (e.g. \cite{1999MNRAS.303..521M,2012AJ....143...72M,2014ApJ...788...41D,2015ApJS..220...14K}).}
Two key arguments that favour circumstellar dust as the cause are i) ``blueing'', where
the star is reddened when moderately dimmed but returns to the stellar colour for very
deep ($\gtrsim$1mag) events -- the remaining flux is interpreted as light scattered off
circumstellar dust \cite{1988SvAL...14...27G}, and ii) increased polarisation fraction
during dimming events, caused by a greater fraction of the flux being contributed by
dust-scattered light (e.g. \cite{2001ARep...45...51R}, which also shows that the
surrounding dust does not reside in a spherical shell). UXors therefore reveal
information on the degree of non-axisymmetry, the ``clumpiness'', of dust orbiting a star
on spatial scales similar to the star itself. The observations can span multiple orbits
to test for repeated dimming events (e.g. \cite{1999A&A...349..619B}), and by using
different bandpasses and polarisation can estimate dust grain sizes
(e.g. \cite{1988SvAL...14...27G,1994ASPC...62...63G}).

In the majority of UXor-like cases (i.e. those related to obscuration by dust), including
other classes such as ``dippers'' (e.g. \cite{2014AJ....147...82C,2016ApJ...816...69A}),
the processes causing young stars to vary are attributed to gas-rich protoplanetary
disks. For Herbig Ae/Be stars the obscuration is thought to be caused by hydrodynamic
turbulence that lifts dust above the puffed up inner rim of a self-shadowed disk
\cite{2003ApJ...594L..47D}. For the dippers, which are observed around low-mass stars,
the obscuration is attributed to dust in accretion streams that link the inner disk and
the stellar surface, and/or to variations in the height of the inner disk edge
\cite{1999A&A...349..619B,2016ApJ...816...69A,2016arXiv160503985B}. The common theme is
therefore that the location of the occulting dust is as close to the star as physically
possible, being set by sublimation (e.g. \cite{2007prpl.conf..539M}). These systems tell
us about the nature of turbulence and accretion in gas-dominated disks, but so far reveal
little about how these disks transition to the debris phase and the subsequent evolution.

Here we focus on RZ~Psc, a star that shows UXor-like variability
(e.g. \cite{1985PZ.....22..181Z,1999A&AS..140..293G,2003ARep...47..580S}). As a young K0V
type star with no evidence for gas accretion and a strong infrared (IR) excess, this
system appears unique among UXors and may provide new information on the structure of
inner planetary systems during or following dispersal of the gas disk
\cite{2010A&A...524A...8G,2013A&A...553L...1D}. However, the IR excess indicates the over
5\% of the starlight is intercepted by the disk, which is a level more akin to gas-rich
protoplanetary and transition disks than debris disks. Specifically, we use a decade of
ground-based optical photometry of RZ~Psc (section \ref{s:data}) to draw conclusions on
dust location (section \ref{s:where}), and discuss the possible disk structure and
evolutionary state in section \ref{s:disk}. We conclude in section \ref{s:conc}.

@@

There is significant evidence that the optical variations seen towards RZ~Psc are caused
by circumstellar dust: i) during dimming events the colour becomes redder
\cite{1985PZ.....22..181Z,1980PZ.....21..310K} in a way consistent with that expected for
dust \cite{1981Afz....17...87P,2004ARep...48..470P}, ii) the maximum depth is about
2.5mag and during these events the colour returns to near stellar values, suggesting that
the remaining emission is from light scattered off the circumstellar dust (i.e. the star
is fully occulted \cite{1981Afz....17...87P,1988SvAL...14...27G}, and iii) the
polarisation fraction increases during the transits, as expected if an increasing
fraction of the light is scattered off a disk of circumstellar dust
\cite{1988SvAL...14...27G,1991Afz....34..333K,2003ARep...47..580S}.

What separates RZ~Psc from other UXors (and dippers) is i) the spectral type is K0V
rather than Herbig Ae/Be for UXors and late K to M-type for dippers, ii) the occulting
dust lies well beyond the sublimation radius, and iii) the star is not associated with a
star-forming region so is inferred to be a few tens of millions of years old. The dust
distance has been inferred from the speed of ingress of dimming events, which was
previously estimated as about 0.6au (for circular orbits
\cite{2013A&A...553L...1D}). Corroborating evidence comes from the $\sim$500K temperature
of the dust seen in the mid-IR, which places it near 0.4-0.7au (depending on optical
depth) and therefore at a location consistent with the occulting dust
\cite{2013A&A...553L...1D}. The distance to RZ~Psc is unknown, but as an apparently
isolated star that shows Li absorption the age has been estimated as a few tens of Myr,
and therefore beyond the age at which a gas-rich disk would normally exist
\cite{2010A&A...524A...8G,2014A&A...563A.139P}. Further distinguishing features are that
the duration of the dimming events is consistently short compared to other UXors, a few
days rather than days to a few weeks, and that no near-IR (i.e. K-band) excess or
accretion signatures are seen \cite{2014A&A...563A.139P}, so interpretations related to
accretion of disk material onto the star
(e.g. \cite{1999AJ....118.1043H,1999A&A...349..619B,2016arXiv160503985B}) are unlikely.

@@

While this asteroid belt picture is intriguing, and makes RZ~Psc a system that could be
of great interest and worthy of detailed study, it is not the only possibility. Well over
1\% of the starlight is reprocessed by the circumstellar disk, which is more typical of
the primordial gas-rich disks seen around nearly all young stars. The discovery of
systems like HD~21997, that appear to be a few tens of Myr old and host gas-rich disks
\cite{2013ApJ...776...77K}, shows that stellar age is not a perfect indicator of disk
status. Thus, a considerable part of our analysis focuses on the question of the status
of the disk around RZ~Psc.

@@

\subsection{Optical}\label{ss:opt}

% yearly-lc.pro
\begin{figure*}
  \begin{center}
    \hspace{-0.5cm} \includegraphics[width=\textwidth]{figs/yearly-2004.eps}
    \caption{WASP and KELT-North data. Photometry is shown in dimensionless form,
      relative to a quiescent level of 1, and was converted from observed magnitudes as
      described in the text.}\label{fig:waspkelt}
  \end{center}
\end{figure*}

To study the temporal variability of RZ~Psc we use two seasons of public data from the
Wide-Angle Search for Planets (WASP \cite{2006PASP..118.1407P,2010A&A...520L..10B}), and
nine seasons of data from the Kilodegree Extremely Little Telescope North (KELT-North
\cite{2007PASP..119..923P}). We also collected, but ultimately did not use, photometric
observations of RZ~Psc from a wide variety of other sources
(\cite{1994AJ....108.1906H,1973IBVS..783....1K,1980PZ.....21..310K,1985PZ.....22..181Z,1991Afz....34..333K,1997AcA....47..467P,2014Ap.....57..491P},
the Catalina Sky Survey, the American Association of Variable Star Observers, the All-Sky
Automated Survey). Aside from the Harvard plate photometry published by G\"urtler et
al. \cite{1999A&AS..140..293G} we have not sought unpublished photometry so the light
curve remains incomplete.\footnote{Normalised photometry is available at
  \href{https://github.com/drgmk/rzpsc}{https://github.com/drgmk/rzpsc}}

Here we focus on the WASP and KELT-North data, as it has not been previously analysed and
has considerably higher cadence (many measurements per night) and temporal coverage
(nightly, weather permitting) than other data sets. The WASP data from 2004 and 2006 are
public and were obtained from an online
archive\footnote{\href{http://wasp.cerit-sc.cz}{http://wasp.cerit-sc.cz}}. These data
were processed in a manner similar to that described by \cite{2014MNRAS.441.2845V}, where
common-mode variations were removed using 50 quiet nearby stars. The WASP bandpass is
broad, with roughly uniform transmission from 400-700nm \cite{2006PASP..118.1407P}. The
KELT-North data, 2006-2014, were used in raw form, the only specific treatment being a
4\% relative correction being made to ensure observations taken in the ``east'' and
``west'' telescope orientations have the same calibration. The bandpass is redder than
for WASP, with most transmission between 500nm and 800nm \cite{2007PASP..119..923P}.  For
a full description of the KELT-North data reduction, see \cite{2012ApJ...761..123S}.

We normalised each year's data from each instrument separately by converting magnitudes
to flux density and dividing out the sigma-clipped median so the light curve has an
out-of-occultation baseline of 1. In doing so we are assuming that variations due to the
slightly different filter bandpasses are unimportant. Each row in Figure
\ref{fig:waspkelt} shows a season's data, starting on May 1 each year (JD also
indicated). Most year's data therefore extend into the next year, so the ``2006 data''
refers to data from the 2006/2007 observing season.

\subsubsection{Qualitative light curve overview}\label{sss:quallc}

@@

It is clear from Figure \ref{fig:waspkelt} that RZ~Psc undergoes the very deep dimming
events that are typical of UXors. These are seen a few times each observing season and
vary in complexity, with a few extended events (e.g. 2006) and a greater number of
``neater'' single events (e.g. 2010). In some years there is also significant variability
at shallower depths. Of particular note is the pair of deep events in 2006; these appear
to be about 70 days apart, and given the suggestion that the putative asteroid belt
analogue resides near 0.5au a natural inference is that these two events are related. If
true, this repetition corresponds to a semi-major axis of about 0.3au, which given
uncertainties in the true disk spectrum could be consistent with the location of the
asteroid belt. In 2004 there are about 100 days of near-consecutive nights of data and no
deep events, so either the true period is longer than 100 days, or dust clumps can be
created (and perhaps destroyed) on timescales of a year or so.

@@

A fairly basic question is whether the light curve could result from objects that all
have the same properties, or whether a range is required. Given the existence of both
long shallow events and short deep events, at a minimum the clumps must vary in size
and/or velocity across the face of the star, but probably also have different optical
depths. The star can be completely occulted, so the clumps can be optically thick and
star-sized, but it has already been shown that the events are not grey in colour, so the
clumps must have a density gradient rather than sharp edges. Where sufficient data exist,
it is clear that not all events have the same relative shape, so the shape of the clumps
varies. Thus, the broad picture is of roughly star-sized clumps, whose shape and orbital
elements vary. The fact that the dimming events can be shown in a series of panels with
the same scale in Figure \ref{fig:waspkeltzoom} suggests that the range over which these
properties vary is of order factors of a few, not many orders of magnitude.

\subsection{Infrared}\label{ss:ir}

While the optical photometry reveals information about how RZ~Psc is itself dimmed, IR
photometry beyond a few microns is dominated by emission from the circumstellar
disk. Thus, IR variation reveals information about how the emitting surface area,
temperature, and perhaps composition, of the dust change with time. Such variation is
indeed apparent, both from comparison of an AKARI 18$\mu$m non-detection at a lower level
than the WISE 22$\mu$m detection, and from several individual WISE measurements taken at
6-month intervals.

% visir analysis, p97-results
\begin{figure*}
  \begin{center}
    \hspace{-0.5cm} \includegraphics[width=0.7\textwidth]{figs/RZ-Psc.eps}
    \caption{Flux density distribution of RZ~Psc, including 2MASS, WISE, AKARI, VISIR,
      and IRAS data and their (approximate) year of observation. The dark blue line shows
      a stellar photosphere model at the approximate stellar temperature of 5350K, and
      the light blue line a 500K blackbody. The latter is not a fit, but an approximate
      continuum level that illustrates that the WISE 3.4 and 22$\mu$m photometry cannot
      both be accounted for with a single blackbody, if the silicate feature seen with
      VISIR was present in 2010.}\label{fig:sed}
  \end{center}
\end{figure*}

Motivated by this variation, we obtained VLT/VISIR observations of RZ~Psc; an N-band
(10$\mu$m) spectrum on 2016 August 16 and Q-band (18$\mu$m) photometry on 2016 July 27
(programme 097.C-0217). These data, and the related calibration observations, were
reduced using the standard ESO \texttt{esorex} pipeline. The wall-clock integration time
for the spectrum was 50 minutes at an airmass of 1.66, and was calibrated using an
observation of HD~189831 taken immediately afterwards at an airmass of 1.63. The spectra
in individual chop/nod cycles are consistent, suggesting that the absolute calibration is
reliable. The spectrum was trimmed to mask highly uncertain regions shortward of 8$\mu$m,
and longward of 13 $\mu$m, and near the telluric absorption at 9.5$\mu$m. The Q-band
photometry took 45 minutes at an airmass of 1.65-1.7, and was calibrated against an
observation of HD~2436 taken immediately afterwards at an airmass of 1.5. Photometry of
RZ~Psc and HD~2436 was done using a 0.9'' radius aperture and a sky annulus from
1-2''. In addition to the conversion from adu/s to Jansky using HD~2436, we applied an
additional upward correction to account for the slightly lower airmass for the calibrator
($\exp(0.3 [1.675-1.5] \approx 1.05$, where an extinction of 0.3 per unit airmass was
assumed). Uncertainties were estimated as the standard deviation of the flux density in
50 apertures around RZ~Psc. The final calibrated flux is $86 \pm 10$ mJy.

The VISIR spectrum and photometry are shown in Figure \ref{fig:sed}, which also shows the
other available near- to far-IR photometry. The absolute level of the spectrum agrees
well with the IRAS, WISE (from the ALLWISE catalogue), and AKARI observations; that four
observations spanning over 30 years are consistent suggests that while this part of the
spectrum may vary, the shape and level shown is typical. The same cannot be said near
18$\mu$m, where the AKARI and VISIR (and perhaps WISE depending on the disk spectrum)
flux densities are inconsistent. The flux near 10$\mu$m being relatively constant, and
the 18$\mu$m flux changing could be indicative of shadowing of an outer disk by an inner
disk; that is, evidence that the disk around RZ~Psc has significant radial extent. This
type of behaviour is seen as ``seesaw'' variability in IR spectra towards some transition
disks (e.g. \cite{2011ApJ...728...49E,2012ApJ...748...71F}).

The spectrum clearly shows solid-state emission (and the continuum level is not actually
clear), which indicates i) that the dust is of order microns in size, and ii) that the
dust includes silicates. The smooth rise and fall suggests that the silicates are largely
amorphous; crystalline silicates have sharper features, notably a depression between two
peaks at 10 and 11 $\mu$m, rather than the flat top seen here. Other systems thought to
host bright asteroid belt analogues (rather than gas-rich disks), such as HD~69830,
BD+20~307, and HD~113766A, tend to show crystalline features
(e.g. \cite{2005ApJ...626.1061B,2005Natur.436..363S,2012A&A...542A..90O}), which may
argue against such a scenario for RZ~Psc. However, such comparisons are largely
speculative as there is also a high degree of variation among silicate features, for both
gas-rich and gas-poor disks.

\begin{figure*}
  \begin{center}
    \hspace{-0.5cm} \includegraphics[width=\textwidth]{figs/wise-var.eps}
    \caption{WISE epoch photometry at 3.4, 4.6, 12, and 22$\mu$m (W1, W2, W3, and W4, in
      magnitudes). The left panel shows the time variation in W1 over 5.5
      years. Subsequent panels show how W1 correlates with W2, W3, and W4, which do not
      have observations at all W1 epochs. The dashed lines show the slope expected for
      constant disk flux variation with wavelength (the variation is smaller in W1/2
      because the total flux is not dominated by the disk).}\label{fig:wisevar}
  \end{center}
\end{figure*}

To further explore this variability, we use the WISE ``epoch'' photometry, for which the
telescope scanning strategy results in clusters of measurements that are spaced 6 months
apart. These data appear at approximately days 70 and 250 in the relevant years in Figure
\ref{fig:waspkelt}, but do not coincide with any optical dimming events. Photometry is
not available in all four channels since launch in early 2010, due to exhaustion of the
coolant after 7.7 months and a 2.5-year hiatus from mid 2011-2014 (see
\cite{2010AJ....140.1868W,2014ApJ...792...30M}). These data are shown in Figure
\ref{fig:wisevar}, where the left panel shows the 3.4$\mu$m magnitude as a function of
time, and that there is significant variation on 6-month timescales, and an even greater
variation overall. Inspection of the individual clusters, which are on hour to day
timescales, shows no significant variation with time. The dashed lines have the slope
expected for disk brightness variation that is independent of wavelength; the slopes are
not exactly 1 because the total flux is not dominated by the disk near 3-5$\mu$m, and
hence the slopes are slightly flatter. Comparing the observed and expected correlations,
we conclude that the data do not show significant evidence for changes in the spectral
shape (i.e. changing temperature or composition). However, the ratios including 12 and
22$\mu$m observations are most sensitive to these change, but only include the first two
sets of observations where the brightness changes were relatively small.

Nevertheless, the amplitude of the change in 3.4$\mu$m brightness over 5 years is about
30\%. Considering that the disk flux density is only 45\% of the total flux at this
wavelength, the disk brightness increased from 2010 to 2015 by about a factor of two. A
similar variation can be inferred by comparing the 18$\mu$m upper limit from AKARI in
2007 and our VISIR measurement in 2016. Given these increases is it surprising that the
N-band spectrum does not appear much higher than the IRAS, AKARI, and WISE photometry. A
possible explanation would be that the increased emission originates in larger grains,
which would result in greater continuum flux but similar levels in spectral
features. However, without wider spectral and more frequent temporal coverage quantifying
such effects is difficult. This level of IR variation is seen towards both protoplanetary
(e.g. \cite{2011ApJ...728...49E,2014ApJ...791...42Z}) and debris disks
(e.g. \cite{2012ApJ...751L..17M,2012Natur.487...74M,2014Sci...345.1032M}), so these data
provide little means to distinguish between scenarios.

@@

The main part of our analysis concerns attempts to extract information from the optical
light curve, taking advantage of the great number of dimming events seen over ten
seasons. In this section we focus on the radial location of the bodies (``clumps'') that
pass in front of RZ~Psc, first searching for periodicity associated with repeat events,
and then using the light curve gradients to constrain the projected velocities. This
analysis primarily focusses on what can be gleaned from the light curves, and the
implications of these results for different clump origins are then explored in section
\ref{s:disk}.

@@

We begin by estimating the lifetime of an occulting clump as a check on the plausibility
that dimming events should repeat. The angular rate at which clumps are sheared
  out is R$d \Omega / dR = - 3 \Omega / 2$. Accounting for shear in both forward
  (interior) and backward (exterior) directions the shear velocity across a clump of
  radius $R_{\rm cl}$ is then
\begin{equation}\label{eq:shear}
  v_{\rm sh} = 3 R_{\rm cl} \Omega \, ,
\end{equation}
so the clump expansion rate due to shear in units of clump radii is only three times the
orbital frequency. That is, after one orbit a clump will be stretched by a factor of
$6\pi$, and the radial and vertical optical depth will be roughly $6\pi$ lower (though it
might still be optically thick). Thus, clumps that are not bound by their own
self-gravity are expected to have a short lifetime at optical depths that are large
enough to cause detectable dimming events, but could cause repeated dimming events if
they are initially optically thick.

@@

% period-search.pro
\begin{figure}
\begin{center}
\hspace{-0.5cm} \includegraphics[width=0.7\textwidth]{figs/auto.eps}
\caption{Discrete autocorrelation function for yearly WASP and KELT-North data, computed
  for lags between 10-155 days). The second and third lines from the top show all data,
  and 2006-excluded data. The topmost line shows the number of years that show a peak
  more than 3$\sigma$ above the clipped DACF mean within each 5 day bin. The peak at 70
  days is 7 years, the dashed line is zero, and the y-axis scale is shown to the
  right.}\label{fig:auto}
\end{center}
\end{figure}
